---
output:
  md_document:
    variant: markdown_github
---

# Purpose


```{r, eval=FALSE}

rm(list = ls()) # Clean your environment:
gc() # garbage collection - It can be useful to call gc after a large object has been removed, as this may prompt R to return memory to the operating system.
library(tidyverse)
list.files('code/', full.names = T, recursive = T) %>% .[grepl('.R', .)] %>% as.list() %>% walk(~source(.))
```

# Loading relevant packages 

```{r setup, eval=FALSE}
install.packages("pacman")
pacman::p_load(tidyverse)
pacman::p_load(readxl)
pacman::p_load(readr)
pacman::p_load(ggplot2)
library(tidyverse)
```


# Question 1 

## Setting up file structure for Question 1 

```{r eval=FALSE}
Texevier::create_template_html(directory = "C:/Users/austi/OneDrive/Desktop/Masters/Data Science/22582053",
                          template_name = "Question 1",
                          build_project = T, open_project = T)

```

```{r}
library(tidyverse)
list.files('C:/Users/austi/OneDrive/Desktop/Masters/Data Science/22582053/Question 1/code', full.names = T, recursive = T) %>% as.list() %>% walk(~source(.))
```

```{r include=FALSE}
covid_data_description <- read_csv("C:/Users/austi/OneDrive/Desktop/Masters/Data Science/22582053/Question 1/data/covid_data_description.csv")
Deaths_by_cause <- read_csv("C:/Users/austi/OneDrive/Desktop/Masters/Data Science/22582053/Question 1/data/Deaths_by_cause.csv")
owid_covid_data <- read_csv("C:/Users/austi/OneDrive/Desktop/Masters/Data Science/22582053/Question 1/data/owid-covid-data.csv")
```

# obtaining the total cases, total deaths on the last date: 

```{r}
Final_total_covid_cases <- obtain_total_covid_casses_function(owid_covid_data)
print(Final_total_covid_cases)

Final_total_covid_deaths <- obtain_total_deaths_function(owid_covid_data)
```

# Comparing the effect of Covid-19 on different continents

## Total cases

```{r}
covid_continent_data <- continent_comparer_function(owid_covid_data, continent) #In this line of code I use a function that groups by continent and then provides a summary for the total cases, total deaths and total new cases. 

bar_plot_covid_data(Final_total_covid_cases)
```

## Total deaths 

```{r}
bar_plot_covid_deaths_data(Final_total_covid_deaths)
```


# Question 2 

## Setting up the file structure for Question 2

```{r eval=FALSE}
Texevier::create_template_html(directory = "C:/Users/austi/OneDrive/Desktop/Masters/Data Science/22582053",
                          template_name = "Question 2",
                          build_project = T, open_project = T)
```

#  QUestion 3 

## Setting up the file structure for Question 3

```{r eval=FALSE}
Texevier::create_template(directory = "C:/Users/austi/OneDrive/Desktop/Masters/Data Science/22582053",
                          template_name = "Question 3",
                          build_project = T, open_project = T)
```

## Data 

The data used is obtained from Spotify for the two bands of Coldplay and Mettallica. The relevant excel files have been placed into the Question 3 data folder and then called from there into my enviroment. 

```{r}
#This code is too load my data into my enviroment
library(readr)
Coldplay <- read_csv("C:/Users/austi/OneDrive/Desktop/Masters/Data Science/22582053/Question 3/data/Coldplay.csv")

Metallica <- read_csv("C:/Users/austi/OneDrive/Desktop/Masters/Data Science/22582053/Question 3/data/metallica.csv")
```

```{r}
library(tidyverse)
list.files('C:/Users/austi/OneDrive/Desktop/Masters/Data Science/22582053/Question 3/code', full.names = T, recursive = T) %>% as.list() %>% walk(~source(.))

#This code is used to call in the code used for Question 3. 
```

## Introduction

This comparison report will take a deep dive into the various attributes of Coldplay and Mettallica, such as, tempo, loudness and valence. this comparison report attempts to understand if their are certain attributes that ensure a bands success. Throughout this report it is found that the two bands are very different and that there are no certainties when it comes to what attributes a band needs to become successful.

# Graphical representation of the data

## Histogram: Plotting histograms to Compare the distribution of tempo values for Coldplay and Metallica songs.

```{r}

# This code uses a function to create a histogram comparing tempo values. 
tempo_distribution_function(Coldplay, Metallica, x_var = "tempo")

```

As can be seen in the above figure, Mettallica has on average a higher tempo count then that of Coldplay. What this means is that Mettallica tends to play songs that are of  a faster pace than Coldplay. What can be taken from this output is that having a fast or slow tempo on average for a band will not hint towards the band not performing well over generations. This is because the above figure is evidence towards two successful bands with differing tempos. 

## Violin Plot: Compare the distribution of loudness values for Coldplay and Metallica songs using a violin plot.

```{r}

#This code uses a function to compare loudness values
data_loudness <- loudness_data_combined(Coldplay, Metallica, "Coldplay", "Metallica", "loudness")

loudness_comparison_plot(data_loudness, x_var = "Band", y_var = "Loudness")

```

The purpose of the violin plot is too see the distribution of values. In the case of the above figure we are analyzing the distribution of the loudness variable for both Coldplay and Mettallica so as to compare how loud the two bands on average play their music. As can be seen from the above figure The mean values for both bands seem to be similar however, the differences come in when we look at the tails. When evaluating the tails of the above figure it is evident that Coldplay tends to the lower volume while Mettallic on the occasion will play the louder music. 

## Box Plot: Compare the distribution of multiple attributes (e.g., danceability, energy, valence) between Coldplay and Metallica songs using box plots side by side.

I tried to use functional coding for this section but then the code wouldnt run properly as i couldnt get it perfect. I therefore settled for the next best thing and went for the ugly looking code. 

```{r}

# Combining attribute data for both bands
attribute_data <- rbind(
  data.frame(Band = rep("Coldplay", nrow(Coldplay)), 
             Danceability = Coldplay$danceability,
             Energy = Coldplay$energy,
             Valence = Coldplay$valence),
  data.frame(Band = rep("Metallica", nrow(Metallica)), 
             Danceability = Metallica$danceability,
             Energy = Metallica$energy,
             Valence = Metallica$valence)
)

# Creating individual boxplots using a fucntion
danceability_boxplot <- attribute_boxplot_creator(attribute_data, "Danceability")

energy_boxplot <- attribute_boxplot_creator(attribute_data, "Energy")

valence_boxplot <-attribute_boxplot_creator(attribute_data, "Valence")

combined_boxplot <- combined_boxplot_creator(attribute_data)

grid.arrange(danceability_boxplot, energy_boxplot, valence_boxplot, combined_boxplot, ncol = 2)

```

The final plot we will be looking at are some combined box plots which aims to explain the differences in the attributes of danceability, energy, and valence. As can be seen by the above figure, Mettallica has a lower danceability when compared to that of Coldplay, however, this is the only measure where Mettallica lags Coldplay. On the other metrics such as, energy and valence, Mettalic out performs Coldplay. 

## Conclusion


Therefore, to conclude, when comparring the two generational bands of Coldplay and Mettallica it is evident that the bands are very different. Mettalica tend to play louder music with more energy, while Coldplay likes to play music that is softer but easier to dance too resulting in the attraction of different crowds. Thus from the output seen, there is no one formula that will lead to a band being successful. A band can be successful being themselves and playing music they enjoy. 

# Question 4 

## Setting up the file structure for Question 4 

```{r eval=FALSE}
Texevier::create_template(directory = "C:/Users/austi/OneDrive/Desktop/Masters/Data Science/22582053",
                          template_name = "Question 4",
                          build_project = T, open_project = T)
```

# Question 5

## Setting up the file structure for Question 5

```{r eval=FALSE}
Texevier::create_template_html(directory = "C:/Users/austi/OneDrive/Desktop/Masters/Data Science/22582053",
                          template_name = "Question 5",
                          build_project = T, open_project = T)
```

