---
# IMPORTANT: Change settings here, but DO NOT change the spacing.
# Remove comments and add values where applicable.
# The descriptions below should be self-explanatory

title: "Question 3: Cold play and Metallica"
#subtitle: "This will appear as Right Header"

documentclass: "elsarticle"

# --------- Thesis title (Optional - set to FALSE by default).
# You can move the details below around as you please.
Thesis_FP: FALSE
# Entry1: "An unbelievable study with a title spanning multiple lines."
# Entry2: "\\textbf{Some Guy}" # textbf for bold
# Entry3: "A thesis submitted toward the degree of Doctor of Philosophy"
# Uni_Logo: Tex/Logo.png # Place a logo in the indicated location (from your root, e.g. defaults to ~/Tex/Logo.png) and uncomment this line. Leave uncommented for no image
# Logo_width: 0.3 # If using a logo - use this to set width (size) of image
# Entry4: "Under the supervision of: \\vfill Prof. Joe Smith and Dr. Frank Smith"
# Entry5: "Stellenbosch University"
# Entry6: April 2020
# Entry7:
# Entry8:

# --------- Front Page
# Comment: ----- Follow this pattern for up to 5 authors
AddTitle: TRUE # Use FALSE when submitting to peer reviewed platform. This will remove author names.
Author1: "Austin Byrne"  # First Author - note the thanks message displayed as an italic footnote of first page.
Ref1: "Music fenatic, Stellenbosch, South Africa" # First Author's Affiliation
Email1: "22582053\\@sun.ac.za" # First Author's Email address

#Author2: "John Smith"
#Email2: "John\\@gmail.com"
#CommonAffiliation_12: TRUE # If Author 1 and 2 have a common affiliation. Works with _13, _23, etc.

#Author3: "John Doe"
#Email3: "Joe\\@gmail.com"

CorrespAuthor_1: TRUE  # If corresponding author is author 3, e.g., use CorrespAuthor_3: TRUE

# Comment out below to remove both. JEL Codes only given if keywords also given.
#keywords: "Multivariate GARCH \\sep Kalman Filter \\sep Copula" # Use \\sep to separate
#JELCodes: "L250 \\sep L100"

# ----- Manage headers and footers:
#BottomLFooter: $Title$
#BottomCFooter:
#TopLHeader: \leftmark # Adds section name at topleft. Remove comment to add it.
BottomRFooter: "\\footnotesize Page \\thepage" # Add a '#' before this line to remove footer.
addtoprule: TRUE
addfootrule: TRUE               # Use if footers added. Add '#' to remove line.

# --------- page margins:
margin: 2.3 # Sides
bottom: 2 # bottom
top: 2.5 # Top
HardSet_layout: TRUE # Hard-set the spacing of words in your document. This will stop LaTeX squashing text to fit on pages, e.g.
# This is done by hard-setting the spacing dimensions. Set to FALSE if you want LaTeX to optimize this for your paper.

# --------- Line numbers
linenumbers: FALSE # Used when submitting to journal

# ---------- References settings:
# You can download cls format here: https://www.zotero.org/ - simply search for your institution. You can also edit and save cls formats here: https://editor.citationstyles.org/about/
# Hit download, store it in Tex/ folder, and change reference below - easy.
bibliography: Tex/ref.bib       # Do not edit: Keep this naming convention and location.
csl: Tex/harvard-stellenbosch-university.csl # referencing format used.
# By default, the bibliography only displays the cited references. If you want to change this, you can comment out one of the following:
#nocite: '@*' # Add all items in bibliography, whether cited or not
# nocite: |  # add specific references that aren't cited
#  @grinold2000
#  @Someoneelse2010

# ---------- General:
RemovePreprintSubmittedTo: TRUE  # Removes the 'preprint submitted to...' at bottom of titlepage
#Journal: "Journal of Finance"   # Journal that the paper will be submitting to, if RemovePreprintSubmittedTo is set to TRUE.
toc: FALSE                       # Add a table of contents
numbersections: TRUE             # Should sections (and thus figures and tables) be numbered?
fontsize: 11pt                  # Set fontsize
linestretch: 1.2                # Set distance between lines.
link-citations: TRUE            # This creates dynamic links to the papers in reference list.

### Adding additional latex packages:
# header-includes:
#    - \usepackage{colortbl} # Add additional packages here.

output:
  pdf_document:
    keep_tex: TRUE
    template: Tex/TexDefault.txt
    fig_width: 3.5 # Adjust default figure sizes. This can also be done in the chunks of the text.
    fig_height: 3.5
abstract: |
  Coldplay and Metallica are two of the greatest bands to have graced the stage in recent decades. This paper attempts a comparisson of the two bands in order to identify how a band can thrive theoughout generations. In this report it is found that the two bands take different approaches to their music. With varying attributes. What this comparisson report establishes, is that, it is not possible to compare bands and make an assuption od what their success will look like. Two very different bands can survive through generations. 
---

<!-- First: Set your default preferences for chunk options: -->

<!-- If you want a chunk's code to be printed, set echo = TRUE. message = FALSE stops R printing ugly package loading details in your final paper too. I also suggest setting warning = FALSE and checking for warnings in R, else you might find ugly warnings in your paper. -->

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, fig.width = 6, fig.height = 5, fig.pos="H", fig.pos = 'H')
# Note: Include = FALSE implies the code is executed, but not printed in your pdf.
# warning and message = FALSE implies ugly messages and warnings are removed from your pdf.
# These should be picked up when you execute the command chunks (code sections below) in your rmd, not printed in your paper!

# Lets load in example data, and see how this can be stored and later called from your 'data' folder.
if(!require("tidyverse")) install.packages("tidyverse")
library(tidyverse)

```

```{r}
list.files('C:/Users/austi/OneDrive/Desktop/Masters/Data Science/22582053/Question 3/code', full.names = T, recursive = T) %>% as.list() %>% walk(~source(.))
```


<!-- ############################## -->
<!-- # Start Writing here: -->
<!-- ############################## -->

# Introduction \label{Introduction}

This comparison report will take a deep dive into the various attributes of Coldplay and Mettallica, such as, tempo, loudness and valence. this comparison report attempts to understand if their are certain attributes that ensure a bands success. Throughout this report it is found that the two bands are very different and that there are no certainties when it comes to what attributes a band needs to become successful.

#  Data 

The data used is obtained from Spotify for the two bands of Coldplay and Mettallica. 

```{r}
#This code is too load my data into my enviroment
library(readr)
Coldplay <- read_csv("C:/Users/austi/OneDrive/Desktop/Masters/Data Science/22582053/Question 3/data/Coldplay.csv")

Metallica <- read_csv("C:/Users/austi/OneDrive/Desktop/Masters/Data Science/22582053/Question 3/data/metallica.csv")
```

# Graphical representation of the data

## Histogram: Plotting histograms to Compare the distribution of tempo values for Coldplay and Metallica songs.

```{r Figure 1, warning =  FALSE, fig.align = 'center', fig.cap = "Tempo distribution \\label{Figure1}", fig.height = 6, fig.width = 6, dev = 'png'}

# This code uses a function to create a histogram comparing tempo values. 
tempo_distribution_function(Coldplay, Metallica, x_var = "tempo")

```

As can be seen in the above figure \ref{Figure1}, Mettallica has on average a higher tempo count then that of Coldplay. What this means is that Mettallica tends to play songs that are of  a faster pace than Coldplay. What can be taken from this output is that having a fast or slow tempo on average for a band will not hint towards the band not performing well over generations. This is because \ref{Figure1}, is evidence towards two successful bands with differing tempos. 

## Violin Plot: Compare the distribution of loudness values for Coldplay and Metallica songs using a violin plot.

```{r Figure 2, warning =  FALSE, fig.align = 'center', fig.cap = "Loudness distribution \\label{Figure2}", fig.height = 6, fig.width = 6, dev = 'png'}

#This code uses a function to compare loudness values
data_loudness <- loudness_data_combined(Coldplay, Metallica, "Coldplay", "Metallica", "loudness")

loudness_comparison_plot(data_loudness, x_var = "Band", y_var = "Loudness")

```

The purpose of the violin plot is too see the distribution of values. In the case of \ref{Figure2}, we are analyzing the distribution of the loudness variable for both Coldplay and Mettallica so as to compare how loud the two bands on average play their music. As can be seen from \ref{Figure2}, The mean values for both bands seem to be similar however, the differences come in when we look at the tails. When evaluating the tails of the \ref{Figure2} it is evident that Coldplay tends to the lower volume while Mettallic on the occasion will play the louder music. 


## Box Plot: Compare the distribution of multiple attributes (e.g., danceability, energy, valence) between Coldplay and Metallica songs using box plots side by side.

```{r Figure 3, warning =  FALSE, fig.align = 'center', fig.cap = "Attribute distribution \\label{Figure3}", fig.height = 6, fig.width = 6, dev = 'png'}

# Combining attribute data for both bands
attribute_data <- rbind(
  data.frame(Band = rep("Coldplay", nrow(Coldplay)), 
             Danceability = Coldplay$danceability,
             Energy = Coldplay$energy,
             Valence = Coldplay$valence),
  data.frame(Band = rep("Metallica", nrow(Metallica)), 
             Danceability = Metallica$danceability,
             Energy = Metallica$energy,
             Valence = Metallica$valence)
)

# Creating individual boxplots using a fucntion
danceability_boxplot <- attribute_boxplot_creator(attribute_data, "Danceability")

energy_boxplot <- attribute_boxplot_creator(attribute_data, "Energy")

valence_boxplot <-attribute_boxplot_creator(attribute_data, "Valence")

combined_boxplot <- combined_boxplot_creator(attribute_data)

grid.arrange(danceability_boxplot, energy_boxplot, valence_boxplot, combined_boxplot, ncol = 2)

```

The final plot we will be looking at are some combined box plots which aims to explain the differences in the attributes of danceability, energy, and valence. As can be seen by \ref{Figure3}, Mettallica has a lower danceability when compared to that of Coldplay, however, this is the only measure where Mettallica lags Coldplay. On the other metrics such as, energy and valence, Mettalic out performs Coldplay. 

# Conclusion 

Therefore, to conclude, when comparring the two generational bands of Coldplay and Mettallica it is evident that the bands are very different. Mettalica tend to play louder music with more energy, while Coldplay likes to play music that is softer but easier to dance too resulting in the attraction of different crowds. Thus from the output seen, there is no one formula that will lead to a band being successful. A band can be successful being themselves and playing music they enjoy. 

<!-- Make title of bibliography here: -->
<!-- \newpage -->



